---
title: "cosimmrSTAN-vignette"
author: "Emma Govan and Andrew Parnell"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{cosimmrSTAN}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


cosimmrSTAN is an R package built to allow for running of more complex SIMMs through STAN. Users can choose whether to use the Variational Bayes (VB) or the Markov chain Monte Carlo (MCMC) algorithms. Users can include random effects, as well as nested effects in this package.

To begin, install the package
```{r, eval = FALSE}
install_github("emmagovan/cosimmrSTAN")
```
Then load it in
```{r, eval = FALSE}
library(cosimmrSTAN)
```


There are several data sets built into the package. We will first use the alligator data set (from Nifong et al) for illustration.
```{r}
alligator_data = cosimmrSTAN::alligator_data
```

For our first example we will look at Length as a fixed effect. We first load in our data using the `cosimmrSTAN_load` function.
```{r}
Length = alligator_data$length
formula = alligator_data$mixtures ~ Length

in_alli<-cosimmrSTAN_load(formula,
                 source_names = alligator_data$source_names,
                 source_means = alligator_data$source_means,
                 source_sds = alligator_data$source_sds,
                 correction_means = alligator_data$TEF_means,
                 correction_sds = alligator_data$TEF_sds)

```


We can print a summary of the loaded in data
```{r}
print.cosimmrSTAN_input(in_alli)
```


We can create an isospace plot, coloured by covariate `Length`
```{r}
plot(in_alli, colour_by_cov = TRUE, cov_name = "Length")
```


We then run the `cosimmrSTAN_in` object through the `cosimmr_stan` function
```{r}
alli_out = cosimmr_stan(in_alli)
```


We can then examine the output
```{r}
summary(alli_out)
```

```{r}
plot(alli_out, type = c("beta_fixed_histogram", "prop_density"))
```

We can then look at a random effect in Wolves data from 


```{r}
wolves_data = cosimmrSTAN::wolves_data
```

For our first example we will look at Length as a fixed effect. We first load in our data using the `cosimmrSTAN_load` function.
```{r}
pack = wolves_data$pack
formula = wolves_data$y ~ pack

in_wolves<-cosimmrSTAN_load(formula,
                 source_names = wolves_data$source_names,
                 source_means = wolves_data$s_mean,
                 source_sds = wolves_data$s_sd,
                 correction_means = wolves_data$c_mean,
                 correction_sds = wolves_data$c_sd)

```


We can print a summary of the loaded in data
```{r}
print.cosimmrSTAN_input(in_wolves)
```


We can create an isospace plot, coloured by covariate `Length`
```{r}
plot(in_wolves, colour_by_cov = TRUE, cov_name = "pack")
```


We then run the `cosimmrSTAN_in` object through the `cosimmr_stan` function
```{r}
wolves_out = cosimmr_stan(in_wolves)
```


We can then examine the output
```{r}
summary(wolves_out)
```

```{r}
plot(wolves_out, type = c("beta_random_histogram", "prop_density"))
```






